---
title: "Effect of smoking on high-blood pressure"
author: "Robin Aldridge-Sutton"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data in the following table was obtained from a study on smoking and high blood  pressure. The researcher wants to know whether people who smoke are more likely to have  high blood pressure.

```{r}
# Gender, male as reference
g = rep(0:1, each = 12)

# Age group, using 20-24 as reference age group
ag = rbind(0, diag(5), 0, diag(5))[rep(1:12, each = 2), ]
colnames(ag) = paste0(c(35, 50, 60, 70, 80), "-")

# Smoking
s = rep(0:1, 12)

# High-blood pressure
hbp = cbind(c(6, 4, 18, 22, 30, 55, 22, 13, 12, 14, 10, 8, 
              2, 9, 12, 18, 28, 22, 24, 32, 22, 41, 11, 4),
            c(18, 11, 28, 17, 22, 44, 30, 52, 32, 26, 22, 5,
              14, 12, 27, 11, 40, 37, 38, 28, 39, 35, 19, 6))
colnames(hbp) = c("yes", "no")
cbind(g, ag, s, hbp)
```
```{r}
# Fit binomial regression model with all interactions
brmais = glm(hbp ~ ag * g * s, family = binomial)
summary(brmais)
```
When all interactions between smoking and the other predictor variables are included there are no statistically significant effects of smoking, so remove the second-order interactions and check each of the first-order interactions.

```{r}
# Fit binomial regression model with interaction between smoking and age-group
brmisag = glm(hbp ~ g + ag * s, family = binomial)
summary(brmisag)
```
The interaction between smoking and age-group not statistically significant.

```{r}
# Fit binomial regression model with interaction between smoking and gender
brmisg = glm(hbp ~ ag + g * s, family = binomial)
summary(brmisg)
```
```{r}
# Standard error, log-odds ratio, and confidence interval
se = sqrt(0.18225^2 + 0.25558^2)
lor = 0.08771 + 0.50865
exp(lor + c(-1, 0, 1) * 1.96 * se)
```
The interaction between gender and smoking is statistically significant, p = 0.047. The effect of smoking for males is not statistically significant, p = 0.63. For females the odds of high blood-pressure are estimate to be 1.8 times higher for smokers than non-smokers, but this is not quite statistically significant (95% confidence interval 0.98, 3.36).
